\vspace*{-8mm}\textbf{\textsc{Novelty Detection for Semantic Place Categorization}}

### 2011-06-26 thesis/novelty-intro.tex
Final fixes.

Other schemes may be used to approximate the unconditional independent assuming that only
a specific variable is not necessarily known. For example modelling unconditional probability
with an hidden variable trained from the unlabelled data. Or by trying to split the single big
factor on factorizations of other factors.
Though the presented cases have interesting properties:

\begin{description}
\item[uniform model] - by assuming the factor to be uniform, it can be considered as not existent on
the graphical model.

\item[independent model] - due to the use of single connected factors, it can easily be trained from
                           unlabelled data. This helps the system to account for an existent bias on
                           each of the variables without allowing the system to overfit to the unlabelled
                           data.
\end{description}

@@

\caption{\label{fig:synthetic-roc}ROC curve comparing novelty detection performance
         under samples with variable size of sensed properties.}
\end{figure}

The convex shape for the optimal threshold shows that the ratio between conditional
and unconditional probability is indeed an optimal detector and is suitable for
implementing a threshold when the samples are taken from dynamic
distributions (e.g.\ some samples where there is only access to room size versus
samples where there is a lot of information about the room properties).

% Discuss importance on approximating unconditional probability.
Its also possible to see how important it is to estimate a correct unconditional
probability in order to obtain a correct novelty measure on the inputs.
The assumption of a uniform unconditional probability has led to very poor results.
That is probably explained by the semantic properties being highly
biased towards some values. And shows that bias plays an important step
in detecting whether a given sensed value is a valuable cue about the room category.

%%% Results 2
% Measure performance of the thresholds as more information becomes available.
\subsection{Performance Changes With Amount of Available Information}
In order to measure the performance impact as more semantic information becomes
available ROC curves were plotted for samples grouped by the number of sensed
semantic features.

@@

\caption{\label{fig:synthetic-roc-breakdown}ROC curves plotted showing performance of the
         presented novelty detection method on graphs generated for different amount of
         sensed features.}
\end{figure}

It is possible to see that as the system gains more semantic information it
becomes easier to detect novelty. The input space size increases and allows the
several existing classes to become more easily distinguished.

@@

The uniform threshold shows a poor performance specially on small size samples
where its performs almost no better than random.
It performance increases as the size of sensed features increases but nonetheless
its very small when compared to how optimal a threshold could be.

>>>>

Other approaches may be used to approximate the single factor described above
and allow modeling the distribution when there is no knowledge about all the states
of a specific variable. For example: by training a hidden variable in an
unsupervised fashion or by using other factorization schemes for the factor.
The presented cases have interesting properties:

\begin{description}
\item[uniform model] - by assuming the factor to be uniform, it can be
considered as not existent in the graphical model.

\item[independent model] - due to the use of single connected factors,
it can easily be trained from unlabelled data. This helps the system to
account for existing bias for each of the variables without the
system overfitting to the unlabelled data.
\end{description}

@@

\caption{\label{fig:synthetic-roc}ROC curve comparing the novelty detection performance
         for samples with variable size of sensed properties.}
\end{figure}

The convex shape of the optimal threshold shows that the ratio between conditional
and unconditional probability is indeed an optimal detector \textbf{[Why does it show it's optimal?? Maybe too strong word?]}
and is suitable for implementing a threshold when the samples are taken from dynamic
distributions (e.g.\ some samples where there is only access to room size versus
samples where there is a lot of information about the room properties).

% Discuss importance on approximating unconditional probability.
Its also possible to see how important it is to estimate a correct unconditional
probability in order to obtain a correct novelty measure on the inputs.
The assumption of a uniform unconditional probability has led to very poor results.
That is probably explained by the semantic properties being highly
biased towards some values. This shows that bias plays an important role
in detecting whether a given sensed value is a valuable cue about the room category.

%%% Results 2
% Measure performance of the thresholds as more information becomes available.
\subsection{Influence of the Amount of Available Information}
In order to measure the performance impact as more semantic information becomes
available, ROC curves were plotted for samples grouped by the number of sensed
semantic features.

@@

\caption{\label{fig:synthetic-roc-breakdown}ROC curves plotted showing performance of the
         presented novelty detection method for graphs generated for different amount of
         sensed features.}
\end{figure}

It is possible to see that as the system gains more semantic information, it
becomes easier to detect novelty. The size of the input space increases and allows the
existing classes to become more easily distinguished.

@@

The uniform threshold shows poor performance especially for samples with small amount of features
where it performs almost no better than random.
The performance increases as the size of sensed features increases but nonetheless
is very small when compared to the optimal threshold.
